Week 1 Quiz
===========

The due date for this quiz is Sun 13 Apr 2014 4:30 PM PDT.


Fix URL reading for knitr. See [Stackoverflow](http://stackoverflow.com/a/20003380).

```{r}
setInternet2(TRUE)
```


Question 1
----------
The American Community Survey distributes downloadable data about United States communities. Download the 2006 microdata survey about housing for the state of Idaho using download.file() from here: 

[https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06hid.csv](https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06hid.csv)

and load the data into R. The code book, describing the variable names is here: 

[https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FPUMSDataDict06.pdf](https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FPUMSDataDict06.pdf)

How many housing units in this survey were worth more than $1,000,000?

```{r}
url <- "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FPUMSDataDict06.pdf"
f <- file.path(getwd(), "PUMSDataDict06.pdf")
download.file(url, f)
url <- "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06hid.csv"
f <- file.path(getwd(), "ss06pid.csv")
download.file(url, f)
require(data.table)
dt <- data.table(read.csv(f))
setkey(dt, VAL)
dt[, .N, key(dt)]
```


Question 2
----------
Use the data you loaded from Question 1. Consider the variable FES in the code book. Which of the "tidy data" principles does this variable violate?

```{r}
setkey(dt, FES)
dt[, .N, key(dt)]
```


Question 3
----------
Download the Excel spreadsheet on Natural Gas Aquisition Program here: 

[https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FDATA.gov_NGAP.xlsx](https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FDATA.gov_NGAP.xlsx)

Read rows 18-23 and columns 7-15 into R and assign the result to a variable called:

`dat`

What is the value of:

`sum(dat$Zip*dat$Ext,na.rm=T)`

(original data source: [http://catalog.data.gov/dataset/natural-gas-acquisition-program](http://catalog.data.gov/dataset/natural-gas-acquisition-program))

```{r}
url <- "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FDATA.gov_NGAP.xlsx"
f <- file.path(getwd(), "DATA.gov_NGAP.xlsx")
download.file(url, f)
require(xlsx)
rows <- 18:23
cols <- 7:15
dat <- read.xlsx(f, colIndex=cols, rowIndex=rows)
sum(dat$Zip*dat$Ext,na.rm=T) 
```


Question 5
----------
The American Community Survey distributes downloadable data about United States communities. Download the 2006 microdata survey about housing for the state of Idaho using download.file() from here: 

https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06pid.csv 

using the `fread()` command load the data into an R object

`DT` 

Which of the following is the fastest way to calculate the average value of the variable

`pwgtp15`

broken down by sex using the `data.table` package?

```{r}
url <- "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06pid.csv"
f <- file.path(getwd(), "ss06pid.csv")
download.file(url, f)
DT <- fread(f)
system.time(sapply(split(DT$pwgtp15,DT$SEX),mean))
system.time(mean(DT$pwgtp15,by=DT$SEX))
system.time(DT[,mean(pwgtp15),by=SEX])
system.time(rowMeans(DT)[DT$SEX==1]) + system.time(rowMeans(DT)[DT$SEX==2])
system.time(mean(DT[DT$SEX==1,]$pwgtp15)) + system.time(mean(DT[DT$SEX==2,]$pwgtp15))
system.time(tapply(DT$pwgtp15,DT$SEX,mean))
```